A modern, event-based architecture for distributed evolutionary algorithms

### 2018-04-03 paper.tex
Reducing space, still to go
Took down to 12 pages with the usual dirty tricks

  This kind of internet-native applications later on transitioned to
  using Service-Oriented Architectures (SoA) \cite{Papazoglou2007}. While monolithic, that is,
  including all services in a single computing node and application, SoA
  were better adapted to heterogeneous environments by distributing
  services across a network using standard protocols. Several authors
  implemented evolutionary algorithms over them
  \cite{garcia2013service,munawar2010design,6955331}. However, scaling
  problems and the extension of cloud deployment and services had made
  this kind of architectures decline in popularity
  \cite{Varghese2018849}.

@@

  % Moved here to make it go to the top of page 3
  \begin{figure*}[t!bp]
  \includegraphics[width=0.95\textwidth]{img/kafka.png}
  \caption{Chart showing the general picture of the layers of a
    serverless architecture, including the messages and services that
    constitute KafkEO, with labels indicating message
    routes and the software components used for every part.}
  \label{fig:kafkeo}
  \end{figure*}
  %
  \section{Event-based architectures and implementing evolutionary
    algorithms over them}
  \label{sec:methods}
  %
  \begin{figure*}[h!tbp]
  \includegraphics[width=0.95\textwidth]{img/kafkEO.png}
  \caption{A flow diagram of KafkEO, showing message routes, MessageHub
    topics and the functions that are being used.}
  \label{fig:kafkeo2}
  \end{figure*}
  % Talked about this earlier: - M
  %The creation of virtualization and isolation mechanisms at many
  %different levels and the consequent granularization of architectures
  %has led to systems that are asynchronous and loosely coupled in what
  %is generally called a microservice architecture.

@@

  \section{Experiments and results}
  \label{sec:res}
  In this section we compare the performance of KafkEO against an implementation of
  the EvoSpace \cite{GValdez2015} pool-based architecture, using the first five
  functions of the Noiseless BBOB testbed \cite{hansen2016coco}, 
  which are real-parameter, single-objective, separable functions, namely: \begin{itemize}
  \item Sphere,
  \item ellipsoidal, which is highly multimodal
  \item Rastrigin,
  \item Buche-Rastrigin,
  \item and the purely lineal function called linear
    slope.
  \end{itemize}

    It is expected that the two algorithms achieve similar results as
  they are functionally equivalent. The EvoSpace implementation follows the
  basic EvoSpace model in which EvoWorkers asynchronously interact with the
  population pool by taking samples of the population to perform a standard
  evolutionary search on the samples, to then return newly evolved solutions back
  to the pool.

@@

    % ALL THIS IS REPEATED TWO PARAGRAPHS ABOVE - Mario
    % no need for it, then - JJ
  %We used an already published algorithm, EvoWorker, as a base for
  %comparison, and used the well known BBOB benchmarks. EvoWorker is
  %based in the EvoSpace \cite{DBLP:conf/evoW/ValdezTVGO13} pool-based
  %evolutionary algorithm, the closest functionally to the stream or
  %event-based EA we are introducing in this paper; the parameters used
  %in this algorithm are shown in Table \ref{tab:params:evoworkers}. As
  %the difficulty of the problem increases, this is reflected in the
  %table parameters by increasing the sample size and the number of
  %workers performing at the same time.
  % Maybe you could explain this algorithm a little bit more.
  % Are we using all these combinations? - JJ
  %% Got mixed up, but it was explained a few paragraphs before.
  \begin{table}
    \small
    \caption{EvoWorker setup, including the number of dimensions of the
      benchmark functions, sample size, which is the size of the block
      of individuals taken and dropped back in the pool, as well as the
      number of actual workers used and samples per worker. }
    \label{tab:params:evoworkers}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension & 2 & 3 & 5 & 10 & 20 & 40\\ \hline
      Iterations per Sample  & 50 & 50 & 50 & 50 & 50 & 50\\ \hline
      Sample Size  & 100 & 100 & 100 & 200 & 200 & 200 \\ \hline
      Samples per Worker & 20 & 30 & 25 & 25 & 25 & 25  \\ \hline
      GA Workers & 2 & 2 & 4 & 5 & 8 & 16  \\ \hline
      Pool Size & 250 & 250 & 500 & 1000 & 2000 & 4000  \\ \hline
    \end{tabular}
  \end{table}

@@

  The evolutionary algorithm implemented in KafkEO used the same code,
  also delegating the evolutionary operations to the standard DEAP
  library, written in
  Python \cite{fortin2012deap}. The parameters for the evolutionary algorithm were fixed and
  are shown in Table \ref{tab:GAparams}. These parameters have been
  chosen as the usual, and default, setup in any experiment, without
  making any attempt to adapt it to this particular flavor of the
  algorithm. In particular, the tournament size injects a high selective
  pressure which is known to decrease diversity. The system also allows
  to set different parameters for every instance; in this proof of concept
  only two parameters were randomly set, {\em Mutation Probability}
  uniformly random in the $[.1,.6]$ range, and {\em Crossover
    Probability} random on $[.8,1]$. This is one deviation from the
  standard evolutionary algorithm, but has been proved in the past to
  provide good results without needing to fine tune different
  parameters \cite{tanabe2013evaluation}.

@@

  %
  \begin{table}
    \small
    \caption{ DEAP GA EvoWorker Parameters }
    \label{tab:GAparams}
    \centering
    \small
    \begin{tabular}{|l|c|}
      \hline
      Selection & Tournament size=12\\ \hline
      Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05  \\ \hline
      Mutation Probability & [.1,.6]  \\ \hline
      Crossover & Two Point  \\ \hline
      Crossover Probability& [.8,1]  \\ \hline
    \end{tabular}
  \end{table}
  %
  %
  \begin{figure*}
  \begin{tabular}{l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l}
  \includegraphics[width=0.266\textwidth]{ppfigdim_f001}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f006}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f002}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f007}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f003}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f008}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f004}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f009}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f005}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f010}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{\label{fig:aRTgraphs}
   \bbobppfigdimlegend{$f_1$ and $f_{24}$}
  }
  \end{figure*}
  %
  The results of the comparison are shown in Figure \ref{fig:aRTgraphs},
  which follow the classical BBOB 2009 format, which includes the amount
  of effort devoted to finding a certain fitness level and time needed
  to do it. The EvoSpace and KafkEO results are shown side by side, only
  for the sake of comparison, since we are only interested for the time
  being in the baseline performance of the proof of concept.

@@

  Other changes will go in the direction of testing the performance of
  the system and computing the cost, so that we can increase the former
  without increasing the latter. Since there is room for increasing
  parallelism, we will try different ways of obtaining better
  algorithmic results by making a parameter sensitivity analysis,
  including population size, length of evolution runs, and other
  algorithmic parameters.

  Once those algorithmic baselines have been set, we will experiment
  with different metaheuristics such as particle swarm optimization, or
  even try for heterogeneous functions with different evolutionary
  algorithm parameters, with the purpose of reducing the number of
  parameters to set at the start.

  On the implementation side, we will work on the automatic deployment
  of this architecture to any free or vendor serverless implementation,
  including maybe abstraction layers such as the {\tt serverless}
  library.

\section*{Acknowledgments}
    The author would like to acknowledge the support of grants\\
    taking\\
    this\\
    much\\
    space\\

>>>>

  This kind of internet-native applications later on transitioned to
  using Service-Oriented Architectures (SoA) \cite{Papazoglou2007}. While monolithic, that is,
  including all services in a single computing node and application, SoA
  were better adapted to heterogeneous environments by distributing
  services across a network using standard protocols. Several authors
  implemented evolutionary algorithms over them
  \cite{garcia2013service,munawar2010design,6955331}. However, scaling
  problems and the extension of cloud deployment and services had made
  this kind of architectures decline in popularity.
%  \cite{Varghese2018849}. Too big citation

@@

  % Moved here to make it go to the top of page 3
  \begin{figure*}[h!tb]
    \centering
  \includegraphics[width=0.75\textwidth]{img/kafka.png}
  \caption{Chart showing the general picture of the layers of a
    serverless architecture, including the messages and services that
    constitute KafkEO, with labels indicating message
    routes and the software components used for every part.}
  \label{fig:kafkeo}
  \end{figure*}

  \section{Event-based architectures and implementing evolutionary
    algorithms over them}
  \label{sec:methods}
  %
  \begin{figure*}[h!tb]
  \includegraphics[width=0.75\textwidth]{img/kafkEO.png}
  \caption{A flow diagram of KafkEO, showing message routes, MessageHub
    topics and the functions that are being used.}
  \label{fig:kafkeo2}
  \end{figure*}
  % Talked about this earlier: - M
  %The creation of virtualization and isolation mechanisms at many
  %different levels and the consequent granularization of architectures
  %has led to systems that are asynchronous and loosely coupled in what
  %is generally called a microservice architecture.

@@

  \section{Experiments and results}
  \label{sec:res}

  In this section we compare the performance of KafkEO against an implementation of
  the EvoSpace \cite{GValdez2015} pool-based architecture, using the first five
  functions of the Noiseless BBOB testbed \cite{hansen2016coco}, 
  which are real-parameter, single-objective, separable functions, namely: Sphere,,  ellipsoidal, which is highly multimodal,  Rastrigin,  Buche-Rastrigin,  and the purely lineal function called linear  slope. It is expected that the two algorithms achieve similar results as
  they are functionally equivalent. The EvoSpace implementation follows the
  basic EvoSpace model in which EvoWorkers asynchronously interact with the
  population pool by taking samples of the population to perform a standard
  evolutionary search on the samples, to then return newly evolved solutions back
  to the pool.

@@

  \begin{table}
    \small
    \caption{EvoWorker setup parameters,
      % including the number of dimensions of the
      % benchmark functions, sample size, which is the size of the block
      % of individuals taken and dropped back in the pool, as well as the
    %   number of actual workers used and samples per worker.
    }
    \label{tab:params:evoworkers}
    \centering
    \small
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Dimension & 2 & 3 & 5 & 10 & 20 & 40\\ \hline
      Iterations per Sample  & 50 & 50 & 50 & 50 & 50 & 50\\ \hline
      Sample Size  & 100 & 100 & 100 & 200 & 200 & 200 \\ \hline
      Samples per Worker & 20 & 30 & 25 & 25 & 25 & 25  \\ \hline
      GA Workers & 2 & 2 & 4 & 5 & 8 & 16  \\ \hline
      Pool Size & 250 & 250 & 500 & 1000 & 2000 & 4000  \\ \hline
    \end{tabular}
  \end{table}

@@

  The evolutionary algorithm implemented in KafkEO used the same code,
  also delegating the evolutionary operations to the standard DEAP
  library, written in
  Python \cite{fortin2012deap},
%  The parameters for the evolutionary algorithm were fixed and
%  are shown in Table \ref{tab:GAparams}.
  using 12 for tournament size, a gaussian mutation sith sigma = 0.05 and a probability between 0.1 and 0.6, plus two point crossover with probability between 0.8 and 1; these are the default parameters.  
  % These parameters have been
  % chosen as the usual, and default, setup in any experiment, without
  % making any attempt to adapt it to this particular flavor of the
  % algorithm.
  In particular, the tournament size injects a high selective
  pressure which is known to decrease diversity. The system also allows
  to set different parameters for every instance; in this proof of concept
  only two parameters were randomly set, {\em Mutation Probability}
  uniformly random in the $[.1,.6]$ range, and {\em Crossover
    Probability} random on $[.8,1]$. This is one deviation from the
  standard evolutionary algorithm, but has been proved in the past to
  provide good results without needing to fine tune different
  parameters \cite{tanabe2013evaluation}.

@@

  %
  % \begin{table}
  %   \small
  %   \caption{ DEAP GA EvoWorker Parameters }
  %   \label{tab:GAparams}
  %   \centering
  %   \small
  %   \begin{tabular}{|l|c|}
  %     \hline
  %     Selection & Tournament size=12\\ \hline
  %     Mutation & Gaussian $\mu=0.0$, $\sigma=0.5$, indbp=0.05  \\ \hline
  %     Mutation Probability & [.1,.6]  \\ \hline
  %     Crossover & Two Point  \\ \hline
  %     Crossover Probability& [.8,1]  \\ \hline
  %   \end{tabular}
  % \end{table}
  %
  %
  \begin{figure*}[h!tb]
  \begin{tabular}{l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l@{\hspace*{-0.025\textwidth}}l}
  \includegraphics[width=0.266\textwidth]{ppfigdim_f001}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f006}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f002}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f007}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f003}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f008}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f004}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f009}\\
  \includegraphics[width=0.266\textwidth]{ppfigdim_f005}&
  \includegraphics[width=0.266\textwidth]{ppfigdim_f010}\\
  \end{tabular}
  \vspace{-3ex}
   \caption{\label{fig:aRTgraphs}
   \bbobppfigdimlegend{$f_1$ and $f_{24}$}
  }
  \end{figure*}
  %
  The results of the comparison are shown in Figure \ref{fig:aRTgraphs},
  which follow the classical BBOB 2009 format, which includes the amount
  of effort devoted to finding a certain fitness level and time needed
  to do it. The EvoSpace and KafkEO results are shown side by side, only
  for the sake of comparison, since we are only interested for the time
  being in the baseline performance of the proof of concept.

@@

  Other changes will go in the direction of testing the performance of
  the system and computing the cost, so that we can increase the former
  without increasing the latter. Since there is room for increasing
  parallelism, we will try different ways of obtaining better
  algorithmic results by making a parameter sensitivity analysis,
  including population size, length of evolution runs, and other
  algorithmic parameters.  Once those algorithmic baselines have been set, we will experiment
  with different metaheuristics such as particle swarm optimization, or
  even try for heterogeneous functions with different evolutionary
  algorithm parameters, with the purpose of reducing the number of
  parameters to set at the start.

  % On the implementation side, we will work on the automatic deployment
  % of this architecture to any free or vendor serverless implementation,
  % including maybe abstraction layers such as the {\tt serverless}
  % library.

\section*{Acknowledgments}

  This paper has been supported in part by
projects TIN2014-56494-C4-3-P (Spanish Ministry of Economy and
Competitiveness) and DeepBio (TIN2017-85727-C4-2-P)
